---
title: "Week 4-Data 606"
author: "Meaghan Burke"
date: "March 18, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chapter 4 Foundations for Inference

Graded: 4.4, 4.14, 4.24, 4.26, 4.34, 4.40, 4.48



###4.4 Heights of adults. 
Researchers studying anthropometry collected body girth measurements
and skeletal diameter measurements, as well as age, weight, height and gender, for 507 physically
active individuals. The histogram below shows the sample distribution of heights in centimeters.3


(a) What is the point estimate for the average height of active individuals? What about the
median? (See the next page for parts (b)-(e).

- Mean:    171.1 cm
- Median:  170.3 cm

(b) What is the point estimate for the standard deviation of the heights of active individuals?
What about the IQR?

- SD: 9.4
- IQR: Q3- Q1 -> 14

(c) Is a person who is 1m 80cm (180 cm) tall considered unusually tall? And is a person who is
1m 55cm (155cm) considered unusually short? Explain your reasoning.

- 180 cm is not too tall, less than 1 SD above the mean
- 155 cm is pretty short, 1.71 SD below the mean

(d) The researchers take another random sample of physically active individuals. Would you
expect the mean and the standard deviation of this new sample to be the ones given above?
Explain your reasoning

- Similar, but not the same as it is random

(e) The sample means obtained are point estimates for the mean height of all active individuals,
if the sample of individuals is equivalent to a simple random sample. What measure do we
use to quantify the variability of such an estimate (Hint: recall that SDx¯ = p
n )? Compute this quantity using the data from the original sample under the condition that the data are a
simple random sample

Standard Error:
```{r}
#sd/ # of people (507)
9.4/sqrt(507)
```
###4.14 Thanksgiving spending, Part I. 
The 2009 holiday retail season, which kicked off on
November 27, 2009 (the day after Thanksgiving), had been marked by somewhat lower self-reported
consumer spending than was seen during the comparable period in 2008. To get an estimate
of consumer spending, 436 randomly sampled American adults were surveyed. Daily consumer
spending for the six-day period after Thanksgiving, spanning the Black Friday weekend and Cyber
Monday, averaged 84.71. A 95% confidence interval based on this sample is (80.31, 89.11).
Determine whether the following statements are true or false, and explain your reasoning.

(a) We are 95% confident that the average spending of these 436 American adults is between
80.31 and 89.11.

- False, the confidence interval estimates range for the population, not just for the sample size.

(b) This confidence interval is not valid since the distribution of spending in the sample is right
skewed
- False, the central limit theorum works here since the sample size > 30. The shape of the distribution sampled doesnt impact the result.

(c) 95% of random samples have a sample mean between 80.31 and 89.11

- False, this does not represents the sample mean, it’s the population mean.

(d) We are 95% confident that the average spending of all American adults is between 80.31 and
89.11.

- True

(e) A 90% confidence interval would be narrower than the 95% confidence interval since we don’t
need to be as sure about our estimate.

- False 

(f) In order to decrease the margin of error of a 95% confidence interval to a third of what it is
now, we would need to use a sample 3 times larger.

- False, it would need to be 9 X bigger 

(g) The margin of error is 4.4.

- True

###4.24 Gifted children, Part I. 
Researchers investigating characteristics of gifted children collected
data from schools in a large city on a random sample of thirty-six children who were identified
as gifted children soon after they reached the age of four. The following histogram shows the distribution
of the ages (in months) at which these children first counted to 10 successfully. Also
provided are some sample statistics.43

(a) Are conditions for inference satisfied?

-  Yes, large & random. Greater than 30

(b) Suppose you read online that children first count to 10 successfully when they are 32 months
old, on average. Perform a hypothesis test to evaluate if these data provide convincing evidence
that the average age at which gifted children fist count to 10 successfully is less than the general
average of 32 months. Use a significance level of 0.10.

```{r}
#as defined by the graphic
n = 36
mean = 30.69
sd = 4.31

standard_error <- sd / sqrt(n)

standard_error

#find the z score
z = (mean - 32) / standard_error

z

# find the p norm
p=pnorm(z)
p
# When compared with significance level of .1, we can reject null hypothesis.0.0341013< .1
```
(c) Interpret the p-value in context of the hypothesis test and the data.

- Gifted children can count to 10 at an earlier age.


(d) Calculate a 90% confidence interval for the average age at which gifted children first count to
10 successfully.

```{r}
n = 36
mean = 30.69
sd = 4.31
z90= 1.645

lower_ci = mean - z90*sd/sqrt(n)
lower_ci

upper_ci = mean + z90*sd/sqrt(n)
upper_ci
```
(e) Do your results from the hypothesis test and the confidence interval agree? Explain.

- Yes, we see the value of 32 months is outside our CI. We would conclude that 32 months is an unusual event.


###4.26 Gifted children, Part II. 
Exercise 4.24 describes a study on gifted children. In this study,
along with variables on the children, the researchers also collected data on the mother’s and father’s
IQ of the 36 randomly sampled gifted children. The histogram below shows the distribution of
mother’s IQ. Also provided are some sample statistics.

(a) Perform a hypothesis test to evaluate if these data provide convincing evidence that the average
IQ of mothers of gifted children is different than the average IQ for the population at large,
which is 100. Use a significance level of 0.10.

```{r}
n=36
mean = 118.2
sd=6.5

standard_error =sd/sqrt(n)
standard_error

z_score = (mean -100)/standard_error

pnorm(z_score)

```
Reject the null hypothesis, conclusion the avg IQ of mother is diff that population at large. 

(b) Calculate a 90% confidence interval for the average IQ of mothers of gifted children
```{r}
lower_ci_m = mean -z90 *standard_error
lower_ci_m
upper_ci_m = mean +z90 *standard_error
upper_ci_m 
```
(c) Do your results from the hypothesis test and the confidence interval agree? Explain

- Yes, We are 90% confident that the mothers iq fall in the confidence interval between 116.33 and 120 and the confidence interval


###4.34 CLT. 

Define the term “sampling distribution” of the mean, and describe how the shape,
center, and spread of the sampling distribution of the mean change as sample size increases.

-  The distribution of sample means,sampling distribution, based on sample of a fixed size. When the distribution is unimodal & symetricm the sample mean should mirror the population mean. This changes as the population size increases. 


###4.40 CFLBs. 
A manufacturer of compact fluorescent light bulbs advertises that the distribution
of the lifespans of these light bulbs is nearly normal with a mean of 9,000 hours and a standard
deviation of 1,000 hours.

(a) What is the probability that a randomly chosen light bulb lasts more than 10,500 hours?
```{r}
pnorm(10500, 9000, 1000, lower.tail = F)
```


(b) Describe the distribution of the mean lifespan of 15 light bulbs.
```{r}
sd = 1000
n = 15
bulb_15 <- sd/sqrt(n)
bulb_15
```


(c) What is the probability that the mean lifespan of 15 randomly chosen light bulbs is more than
10,500 hours?
```{r}

pnorm(10500, mean=9000, sd=bulb_15, lower.tail = FALSE)
```


(d) Sketch the two distributions (population and sampling) on the same scale.
```{r}
SN <- seq(9000 - (4 * sd), 9000 + (4 * sd), length=15)
SR <- seq(9000 - (4 * bulb_15), 9000 + (4 * bulb_15), length=15)
nd1 <- dnorm(SN,9000,1000)
nd2 <- dnorm(SR,9000,bulb_15)

plot(SN, nd1, type="l",col="blue",
     xlab=" Population vs Sampling", 
     main="Light Bulbs",
     ylim=c(0,0.002))
lines(SR, nd2, col="red")
```


(e) Could you estimate the probabilities from parts (a) and (c) if the lifespans of light bulbs had
a skewed distribution?

-  It would be difficult because the sample size is small


4.48 Same observation, di↵erent sample size. Suppose you conduct a hypothesis test based
on a sample where the sample size is n = 50, and arrive at a p-value of 0.08. You then refer back
to your notes and discover that you made a careless mistake, the sample size should have been
n = 500. Will your p-value increase, decrease, or stay the same? Explain.

- As the sample size increases, the p value will decrease.
