---
title: "Chapter 5 Homework"
author: "Meaghan Burke"
date: "March 25, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Chapter 5 - Inference for Numerical Data
Practice: 5.5, 5.13, 5.19, 5.31, 5.45
Graded: 5.6, 5.14, 5.20, 5.32, 5.48



##5.6 Working backwards, Part II. 

A 90% confidence interval for a population mean is (65,77). 
The population distribution is approximately normal and the population standard deviation
is unknown. This confidence interval is based on a simple random sample of 25 observations.
Calculate the sample mean, the margin of error, and the sample standard deviation

```{r}
n <- 25
#(b-a)/2
margin.error <- ((77-65)/2)
margin.error 
```
```{r}
#(a+b)/2 
smaple.mean <- ((77+65)/2)
smaple.mean 
```
```{r}
#Sample standard devation: ME = t(.05)*s/sqrt(n). Using the qt function and df = 25-1 we get

df <- 25-1
t.value <- qt(.95, df)
t.value
```


```{r}
sd <- (margin.error/t.value)*5
sd
```


##5.14 SAT scores. 

SAT scores of students at an Ivy League college are distributed with a
standard deviation of 250 points. Two statistics students, Raina and Luke, want to estimate the
average SAT score of students at this college as part of a class project. They want their margin of
error to be no more than 25 points.

(a) Raina wants to use a 90% confidence interval. How large a sample should she collect?
```{r}
z <- 1.65 
margin.error <- 25
sd <- 250

n <- ((z * sd) / margin.error) ^ 2
round(n) 
```

(b) Luke wants to use a 99% confidence interval. Without calculating the actual sample size,
determine whether his sample should be larger or smaller than Raina’s, and explain your
reasoning.

Since we are multiplying the SD by a larger number, we will get a larger sample size for Luke’s 99% interval.

(c) Calculate the minimum required sample size for 


```{r}
z <- 2.575 
margin.error <- 25
sd <- 250

n <- ((z * sd) / margin.error ) ^ 2
round(n)
```

##5.20 High School and Beyond, Part I. 
The National Center of Education Statistics conducted a survey of high school seniors, 
collecting test data on reading, writing, and several other subjects.
Here we examine a simple random sample of 200 students from this survey. Side-by-side box plots
of reading and writing scores as well as a histogram of the differences in scores are shown below.

(a) Is there a clear difference in the average reading and writing scores?

I do not see a clear difference in the average reading and writing scores 

(b) Are the reading and writing scores of each student independent of each other?

It appears that they are independent of one another 

(c) Create hypotheses appropriate for the following research question: is there an evident difference
in the average scores of students in the reading and writing exam?

The difference of average in between reading and writing equal zero(μr−μw=0)
The difference of average in between reading and writing does NOT equal zero (μr−μw≠0)

(d) Check the conditions required to complete this test.

independence & nearly normal distribution. The box plot does not indicate any skewness or outliers. The conditions are seemingly satisfied.

(e) The average observed difference in scores is xreadwrite = 0.545, and the standard deviationof the differences is 8.887 points. Do these data provide convincing evidence of a difference between the average scores on the two exams?

```{r}
n <- 200
mean.diff <- -.545
df <- n-1
SD <- 8.887
SE <- SD/sqrt(n)
T <- (mean.diff-0)/SE
pvalue <- pt(T, df)
pvalue
```

(f) What type of error might we have made? Explain what the error means in the context of the
application.

Type 1: Incorrectly reject the null hypothesis.

Type 2: Incorrectly reject the alternative hypothesis.

In the case, we may have made a type 2 error by rejecting the alternative hypothesis. I may have wrongly concluded that there is not a difference in the average student reading and writing exam scores.


(g) Based on the results of this hypothesis test, would you expect a confidence interval for the
average difference between the reading and writing scores to include 0? Explain your reasonin

Yes, I would expect a confidence interval for the average difference between reading and writing scores to include 0.


##5.32 Fuel eciency of manual and automatic cars, Part I. 
Each year the US Environmental Protection Agency (EPA) releases fuel economy data on cars manufactured in that year.
Below are summary statistics on fuel eciency(in miles/gallon) from random samples of cars with
manual and automatic transmissions manufactured in 2012. Do these data provide strong evidence
of a difference between the average fuel eciencyof cars with manual and automatic transmissions
in terms of their average city mileage? Assume that conditions for inference are satisfi

```{r}
n.a <- 26
n.m <- 26
SD.A <- 3.58
SD.M <- 4.51
mean.A <- 16.12
mean.M <- 19.85
alpha <- .05
meandiff <- mean.A - mean.M
SE.A <- SD.A/sqrt(n.a) 
SE.M <- SD.M/sqrt(n.m)
SE <- sqrt(((SE.A)^2)+(SE.M)^2)
T.1 <- (meandiff-0)/SE
pvalue.1 <- pt(T.1, 25)
pvalue.1 <- 2*pvalue.1 #because we are running a two tailed test, we multiply by 2
pvalue.1
```
##5.48 Work hours and education. 
The General Social Survey collects data on demographics,
education, and work, among many other characteristics of US residents.47 Using ANOVA, we can
consider educational attainment levels for all 1,172 respondents at once. Below are the distributions
of hours worked by educational attainment and relevant summary statistics that will be helpful in
carrying out this analysis.


(a) Write hypotheses for evaluating whether the average number of hours worked varies across the
five groups.

1) The difference of all the averages is equal
2) There is one average that is not equal to the other ones.

(b) Check conditions and describe any assumptions you must make to proceed with the test.

1)  observations are independent within and across groups

2) The data within each group are nearly normal

3)The variability across the groups is about equal

(c) Below is part of the output associated with this test. Fill in the empty cells.
```{r}
mean <- c(38.67, 39.6, 41.39, 42.55, 40.85)
sd <- c(15.81, 14.97, 18.1, 13.62, 15.51)
n <- c(121, 546, 97, 253, 155)
data_table <- data.frame (mean , sd, n)
data_table
```
```{r}
n <- sum(data_table$n)
k <- length(data_table$mean)

# Finding degrees of freedom
df <- k - 1
dfResidual <- n - k

```



```{r}
# Compute the SSG
meanTotal <- 40.45
SSG <- sum( data_table$n * (data_table$mean - meanTotal)^2 )
SSG 
# Compute the MSG
MSG <- (1 / df) * SSG
MSG 
# Compute the F statistic
F <- 501.54 / 229.12
F
```

d) What is the conclusion of the test?

Given the p-value = 0.0682 is greater than 0.05, I conclude that there is not a significant difference between the groups and therefore I don't reject the null hypothesis.

