---
title: 'Inference for numerical data'
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

## North Carolina births

In 2004, the state of North Carolina released a large data set containing 
information on births recorded in this state. This data set is useful to 
researchers studying the relation between habits and practices of expectant 
mothers and the birth of their children. We will work with a random sample of 
observations from this data set.

## Exploratory analysis

Load the `nc` data set into our workspace.


```{r setup, cache = F}
knitr::opts_chunk$set(error = TRUE)
```
```{r}

library(dplyr)
library(ggplot2)
```

```{r load-data, eval=TRUE}

load("more/nc.RData")
```

We have observations on 13 different variables, some categorical and some 
numerical. The meaning of each variable is as follows.

variable         | description
---------------- | -----------
`fage`           | father's age in years.
`mage`           | mother's age in years.
`mature`         | maturity status of mother.
`weeks`          | length of pregnancy in weeks.
`premie`         | whether the birth was classified as premature (premie) or full-term.
`visits`         | number of hospital visits during pregnancy.
`marital`        | whether mother is `married` or `not married` at birth.
`gained`         | weight gained by mother during pregnancy in pounds.
`weight`         | weight of the baby at birth in pounds.
`lowbirthweight` | whether baby was classified as low birthweight (`low`) or not (`not low`).
`gender`         | gender of the baby, `female` or `male`.
`habit`          | status of the mother as a `nonsmoker` or a `smoker`.
`whitemom`       | whether mom is `white` or `not white`.

1.  What are the cases in this data set? How many cases are there in our sample?

As a first step in the analysis, we should consider summaries of the data. This
can be done using the `summary` command:

```{r summary, eval=TRUE}
summary(nc)
```

As you review the variable summaries, consider which variables are categorical 
and which are numerical. For numerical variables, are there outliers? If you 
aren't sure or want to take a closer look at the data, make a graph.

Consider the possible relationship between a mother's smoking habit and the 
weight of her baby. Plotting the data is a useful first step because it helps 
us quickly visualize trends, identify strong associations, and develop research
questions.

2.  Make a side-by-side boxplot of `habit` and `weight`. What does the plot 
highlight about the relationship between these two variables?

The box plots show how the medians of the two distributions compare, but we can
also compare the means of the distributions using the following function to 
split the `weight` variable into the `habit` groups, then take the mean of each
using the `mean` function.
```{r}
ggplot(data=nc) + geom_boxplot(aes(x=habit, y=weight)) 
```
```{r}
#the by() function was not working. filtered the population to nonsmoker and calculated the mean
non.smoker<-filter(nc, habit == "nonsmoker")

mean(non.smoker$weight)
```
```{r}
#the by() function was not working. filtered the population to smoker and calculated the mean
smoker<-filter(nc, habit == "smoker")

mean(smoker$weight)
```


There is an observed difference, but is this difference statistically 
significant? In order to answer this question we will conduct a hypothesis test
.

## Inference

3.  Check if the conditions necessary for inference are satisfied. Note that 
you will need to obtain sample sizes to check the conditions. You can compute 
the group size using the same `by` command above but replacing `mean` with 
`length`.

```{r}
#the by() function was not working. filtered the population to nonsmoker and calculated the length
non.smoker<-filter(nc, habit == "nonsmoker")

length(non.smoker$weight)
```

```{r}
#the by() function was not working. filtered the population to smoker and calculated the length
smoker<-filter(nc, habit == "smoker")

length(smoker$weight)
```
```{r}
#Distribution normal?
g <- ggplot(nc, aes(x = weight, colour = habit)) + geom_density()
g
```

4.  Write the hypotheses for testing if the average weights of babies born to 
smoking and non-smoking mothers are different.

Next, we introduce a new function, `inference`, that we will use for conducting
hypothesis tests and constructing confidence intervals. 

###The infernce function was not working for me. calculated the summary statistics in a few steps 
```{r}
#calculated the SD of a non smoker 
non.smoker<-filter(nc, habit == "nonsmoker")

sd(non.smoker$weight)
```

```{r,echo = FALSE, warning = FALSE, message=FALSE, error=TRUE}
inference(y = non.smoker$weight, est = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical")
```

```{r}
#calculated the SD of a smoker 
smoker<-filter(nc, habit == "smoker")

sd(smoker$weight)
```
```{r error=TRUE}
inference(y = smoker$weight, est = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical")
```
Summary statistics:
- n_nonsmoker = 873, mean_nonsmoker = 7.1443, sd_nonsmoker = 1.5187
- n_smoker = 126, mean_smoker = 6.8287, sd_smoker = 1.3862




Let's pause for a moment to go through the arguments of this custom function. 
The first argument is `y`, which is the response variable that we are 
interested in: `nc$weight`. The second argument is the explanatory variable, 
`x`, which is the variable that splits the data into two groups, smokers and 
non-smokers: `nc$habit`. The third argument, `est`, is the parameter we're 
interested in: `"mean"` (other options are `"median"`, or `"proportion"`.) Next
we decide on the `type` of inference we want: a hypothesis test (`"ht"`) or a 
confidence interval (`"ci"`). When performing a hypothesis test, we also need 
to supply the `null` value, which in this case is `0`, since the null 
hypothesis sets the two population means equal to each other. The `alternative`
hypothesis can be `"less"`, `"greater"`, or `"twosided"`. Lastly, the `method` 
of inference can be `"theoretical"` or `"simulation"` based.

5.  Change the `type` argument to `"ci"` to construct and record a confidence 
interval for the difference between the weights of babies born to smoking and 
non-smoking mothers.

By default the function reports an interval for ($\mu_{nonsmoker} - \mu_{smoker}$)
. We can easily change this order by using the `order` argument:
```{r inf-weight-habit-ht, eval=TRUE, tidy=FALSE}
inference(y = non.smoker$weight, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical")
```

```{r inf-weight-habit-ht, eval=TRUE, tidy=FALSE}
inference(y = smoker$weight, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical")
```
NON SMOKER
Summary statistics: mean = 7.1443 ;  sd = 1.5187 ;  n = 873 
Standard error = 0.0514 
95 % Confidence interval = ( 7.0435 , 7.245 )

SMOKER
Summary statistics: mean = 6.8287 ;  sd = 1.3862 ;  n = 126 
Standard error = 0.1235 
95 % Confidence interval = ( 6.5867 , 7.0708 )

* * *


## On your own

-   Calculate a 95% confidence interval for the average length of pregnancies 
(`weeks`) and interpret it in context. Note that since you're doing inference 
on a single population parameter, there is no explanatory variable, so you can 
omit the `x` variable from the function.
```{r}
inference(y = nc$weeks, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical")
```
We are 95% confident that we have captured the mean pregnancy length in weeks of the population between 38.1528 weeks and 38.5165 weeks.



-   Calculate a new confidence interval for the same parameter at the 90% 
confidence level. You can change the confidence level by adding a new argument 
to the function: `conflevel = 0.90`.
```{r}
inference(y = nc$weeks, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical", conflevel = 0.90)
```
We are 90% confident that we have captured the mean pregnancy length in weeks of the population between 38.182 weeks and 38.4873 weeks. Note the difference between the upper and lower boundary is smaller than the 95% CI.


-   Conduct a hypothesis test evaluating whether the average weight gained by 
younger mothers is different than the average weight gained by mature mothers.

Inference function is not working for the two variables, I filtered the dataset for the applicable  subset.
```{r}
#the inference() function was not working. filtered the population to nonsmoker and calculated the length
young.mother<-filter(nc, mature == "younger mom")
inference(y = young.mother$gained, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical")
```
```{r}
#the inference() function was not working. filtered the population to nonsmoker and calculated the length
mature.mother<-filter(nc, mature == "mature mom")
inference(y = mature.mother$gained, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical")
```
Observed difference between means (mature mom-younger mom) = -1.7697
Young Mothers:
Summary statistics: mean = 30.5604 ;  sd = 14.3469 ;  n = 844 
95 % Confidence interval = ( 29.5925 , 31.5283 )
Mature Mothers:
Summary statistics: mean = 28.7907 ;  sd = 13.4824 ;  n = 129 
95 % Confidence interval = ( 26.4641 , 31.1173 )

Accept the Null Hypothesis that there is no difference in mean weight gain of the two populations.


-   Now, a non-inference task: Determine the age cutoff for younger and mature 
mothers. Use a method of your choice, and explain how your method works.
```{r}
summary(nc$mage)
```
```{r}
by(nc$mage,nc$mature, length)
```
```{r}
#We know that 75% of women are below 32 and none above 50. 
ge_32 <- nc$mage >= 33.0 # this will give true if >= 33, false if < 33
table(ge_32) #Table will report the total true and the total false.
```
The summary() function was helpful in determining the range/distribution of the mother's age. The by() with the length operator command was helpful in determining how many mature moms there were. I then tested the mother's age data against a condition that had to be above Q3 which had a cutoff of 32 years. I selected 33 years as my starting point.



-   Pick a pair of numerical and categorical variables and come up with a 
research question evaluating the relationship between these variables. 
Formulate the question in a way that it can be answered using a hypothesis test
and/or a confidence interval. Answer your question using the `inference` 
function, report the statistical results, and also provide an explanation in 
plain language.

```{r}
#the inference() function was not working. filtered the population by gender
male.gender<-filter(nc, gender== "male")


inference(y = male.gender$mage, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical")

```
```{r}
#the inference() function was not working. filtered the population by gender
female.gender<-filter(nc, gender== "female")


inference(y = female.gender$mage, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical")

```
Male Babies
Summary statistics: mean = 26.9457 ;  sd = 6.1823 ;  n = 497 
Standard error = 0.2773 
95 % Confidence interval = ( 26.4021 , 27.4892 )


Female Babies
Summary statistics: mean = 27.0537 ;  sd = 6.25 ;  n = 503 
Standard error = 0.2787 
95 % Confidence interval = ( 26.5075 , 27.5999 )


Observed difference between means= 0.108

<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons 
Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0).
This lab was adapted for OpenIntro by Mine &Ccedil;etinkaya-Rundel from a lab 
written by the faculty and TAs of UCLA Statistics.
</div>
