---
title: "SQL & R- Movie Ratings"
author: "Meaghan Burke"
date: "February 11, 2018"
output: html_document
---

###Choose six recent popular movies.  Ask at least five people that you know (friends, family, classmates, imaginary friends) to rate each of these movie that they have seen on a scale of 1 to 5.  Take the results (observations) and store them in a SQL database.  Load the information into an R dataframe.


###Establish the connection

```{r, message=FALSE}
#Connection to SQL table and relevant libraries 
library(RMySQL)
library(sqldf)
```


###Opening database connection
```{r, message=FALSE}
#Hide actual connection due to password 

#connection <- dbConnect(MySQL(), user='root', password='*******', 
                 #dbname='movies_db', host='127.0.0.1')

#dbListTables(connection)


```
###Fetch the data & print out samples 
```{r, message=FALSE}

#establish the connection and fetch the data from the tables 
imdb_info_rs <- dbSendQuery(connection , "select * from imdb_table")

imdb_info_df <- fetch(imdb_info_rs, n=-1)


```
```{r echo = FALSE, results= 'asis'}
library(knitr)
kable(imdb_info_df[1:5, ], caption = "imdb_info_df" )
```


```{r, message=FALSE}
ratings_rs <- dbSendQuery(connection , "select * from rating_dim")

rating_info_df <- fetch(ratings_rs, n=-1)
```

```{r echo = FALSE, results= 'asis'}

kable(rating_info_df[1:5, ], caption = "rating_info_df" )
```



```{r, message=FALSE}
genre_rs <- dbSendQuery(connection , "select * from genre_dim")

genre_info_df <- fetch(genre_rs, n=-1)


```

```{r echo = FALSE, results= 'asis'}

kable(genre_info_df[1:5, ], caption = "genre_info_df" )
```
###Merge the data on "Title_Generated_Number"
```{r setup, message=FALSE}
comprehensive_rating<- merge(x = imdb_info_df, y = rating_info_df, by = "Title_Generated_Number", all = TRUE)
comprehensive_rating<- merge(x = comprehensive_rating, y = genre_info_df, by = "Title_Generated_Number", all = TRUE)


head(comprehensive_rating,10)
```
###Explore the data
```{r, message=FALSE}
#Display the summary statistics of "comprehensive_ratings""
summary(comprehensive_rating)
```

```{r, message=FALSE}
#Display the column names of "comprehensive_ratings"
names(comprehensive_rating)
```

```{r, message=FALSE}
#Display the structure of "comprehensive_rating".
str(comprehensive_rating)
```
```{r, message=FALSE}
library(ggplot2)
# Basic barplot. Average family rank for the top imdb movies 
p<-ggplot(data=comprehensive_rating, aes(x=Title_Generated_Number, y=Average__Ranking)) +
  geom_bar(stat="identity", fill="steelblue")+
  geom_text(aes(label=Average__Ranking), vjust=-0.3, size=3.5)+
  theme_minimal()
p
   
# Horizontal bar plot
p + coord_flip()
```
```{r, message=FALSE}
# Basic scatter plot with regression line IMDB Rank VS Average family Rank
ggplot(comprehensive_rating, aes(x=Rank_IMBD, y=Average__Ranking)) + geom_point(size=2, shape=23)+
  geom_smooth(method=lm)
```
```{r, message=FALSE}
# Basic barplot. Count the number of films in each genre 
ggplot(data=comprehensive_rating,aes(x=genre)) + geom_bar(fill="darkred")


```
```{r, message=FALSE}
# Basic barplot. Count the number of films in each genre 
ggplot(data=comprehensive_rating,aes(x=Year)) + geom_bar(fill="green")

```

