---
title: "Project 1-Data 607"
author: "Meaghan"
date: "February 25, 2018"
output: html_document
---
## Project 1 Scope

In this project, you’re given a text file with chess tournament results where the information has some structure. Your
job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players:
Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents

```{r, message=FALSE}
library(data.table)
library(stringr)

original_file <- data.frame(read.table('C:/Users/burke/OneDrive/Desktop/Data 607/tournamentinfo_project1.txt', 
									   sep = ",", fill = TRUE ))

reformatted_file <- data.frame(original_file[grep("-----------------------------------------------------------------------------------------", original_file$V1, invert = TRUE) , ])
```







```{r}
#split into oddd and even rows to group the values appropriately
odd_rows <- data.frame(reformatted_file[ !c(TRUE,FALSE), ])  # rows


even_rows <-data.frame(reformatted_file[ !c(FALSE,TRUE), ])  # rows

```


```{r results = 'asis'}
#reformat the evens data 
split_even <- strsplit(as.character(even_rows$reformatted_file..c.FALSE..TRUE....), "|", fixed = TRUE)

Pair <- sapply(split_even , "[", 1)
Player_Name<- sapply(split_even , "[", 2)
Total<- sapply(split_even , "[", 3)
Total <- as.numeric(as.character(Total))
Round_1<- sapply(split_even , "[", 4)
Round_2<- sapply(split_even , "[", 5)
Round_3<- sapply(split_even , "[", 6)
Round_4<- sapply(split_even , "[", 7)
Round_5<- sapply(split_even , "[", 8)
Round_6<- sapply(split_even , "[", 9)
Round_7<- sapply(split_even , "[", 10)



evens_data  <- data.frame (Pair, Player_Name, Total, Round_1, Round_2, Round_3, Round_4,
						   Round_5, Round_6, Round_7)
evens_data <- evens_data[-1 ,]

knitr::kable(head(evens_data,10),  caption = "Initial Even Table")
```

```{r results = 'asis'}
sub_evens  <- evens_data[,c("Round_1","Round_2", "Round_3", "Round_4", "Round_5", "Round_6", "Round_7")]
expression_9 = "[A-Z]"
opp_1  <- str_replace_all(sub_evens$Round_1, expression_9, "")
opp_2  <- str_replace_all(sub_evens$Round_2, expression_9, "")
opp_3  <- str_replace_all(sub_evens$Round_3, expression_9, "")
opp_4  <- str_replace_all(sub_evens$Round_4, expression_9, "")
opp_5  <- str_replace_all(sub_evens$Round_5, expression_9, "")
opp_6  <- str_replace_all(sub_evens$Round_6, expression_9, "")
opp_7  <- str_replace_all(sub_evens$Round_7, expression_9, "")
opp_df <-data.frame (opp_1, opp_2, opp_3, opp_4, opp_5, opp_6, opp_7)

knitr::kable(head(opp_df,10),  caption = " Opponents/Round")
```
```{r results = 'asis'}
#win/loss/draw metadata
sub_evens  <- evens_data[,c("Round_1","Round_2", "Round_3", "Round_4", "Round_5", "Round_6", "Round_7")]
expression_8 = "[:digit:]"
winlossdraw_1  <- str_replace_all(sub_evens$Round_1, expression_8, "")
winlossdraw_2 <- str_replace_all(sub_evens$Round_2, expression_8, "")
winlossdraw_3<- str_replace_all(sub_evens$Round_3, expression_8, "")
winlossdraw_4<- str_replace_all(sub_evens$Round_4, expression_8, "")
winlossdraw_5<- str_replace_all(sub_evens$Round_5, expression_8, "")
winlossdraw_6<- str_replace_all(sub_evens$Round_6, expression_8, "")
winlossdraw_7<- str_replace_all(sub_evens$Round_7, expression_8, "")
winlossdraw_df <-data.frame (winlossdraw_1, winlossdraw_2, winlossdraw_3, winlossdraw_4, winlossdraw_5, winlossdraw_6, winlossdraw_7)

knitr::kable(head(winlossdraw_df,10),  caption = "Win/Loss/Draw")
```
```{r results = 'asis'}
#filter out the important information from the evens dataframe 
reformatted_evens_df <- data.frame(evens_data[,c("Pair", "Player_Name", "Total")],
								   opp_df,
								   winlossdraw_df)

knitr::kable(head(reformatted_evens_df,10),  caption = "Even Rows- Data Table")

```




```{r results = 'asis'}
#reformatting of the odd data 

split_odd <-strsplit(as.character(odd_rows$reformatted_file..c.TRUE..FALSE...), "|", fixed = TRUE)

Num <- sapply(split_odd , "[", 1)
USCF_ID_Rtg_Pre_Post <- sapply(split_odd , "[", 2)
Pts<- sapply(split_odd , "[", 3)
Pts <- str_replace(Pts , "N:", "")
Pts <- as.numeric(as.character(Pts))

color_round_1<- sapply(split_odd , "[", 4)
color_round_2<- sapply(split_odd , "[", 5)
color_round_3<- sapply(split_odd , "[", 6)
color_round_4<- sapply(split_odd, "[", 7)
color_round_5<- sapply(split_odd , "[", 8)
color_round_6<- sapply(split_odd , "[", 9)
color_round_7<- sapply(split_odd , "[", 10)

odds_data <-data.frame (Num, USCF_ID_Rtg_Pre_Post, Pts, color_round_1, color_round_2, color_round_3, 
						color_round_4, color_round_5, color_round_6, color_round_7)
odds_data <- odds_data[-1 ,]


knitr::kable(head(odds_data,10),  caption = "Odd Rows- Data Table")
```




```{r}

#isolate the USCF ID And the pre and post ratings 

test <- data.frame(USCF_ID_Rtg_Pre_Post)
id_split <-strsplit(as.character(test$USCF_ID_Rtg_Pre_Post), "/", fixed = TRUE)
USCF_ID <- sapply(id_split , "[", 1)
USCF_ID <- str_trim(USCF_ID)
head(USCF_ID,10)
```






```{r}
Rtg_Pre_Post <- sapply(id_split , "[", 2)
Rtg_Pre_Post <- data.frame(Rtg_Pre_Post)
names(Rtg_Pre_Post)
ratings_split <-strsplit(as.character(Rtg_Pre_Post$Rtg_Pre_Post), "->", fixed = TRUE)

#Pre Ratings
Pre_ratings <-sapply(ratings_split , "[", 1)
Pre_ratings <- str_replace(Pre_ratings , " R: ", "")
expression_10 = "[A-Z]+\\d{2}"
expression_11 = "[A-Z]+\\d{1}"
Pre_ratings  <- str_replace_all(Pre_ratings, expression_10, "")
Pre_ratings  <- str_replace_all(Pre_ratings, expression_11, "")
Pre_ratings <- str_trim(Pre_ratings)
head(Pre_ratings,10)
```

```{r}
#Post Ratings 

Post_ratings <-sapply(ratings_split , "[", 2)
Post_ratings <- str_replace(Post_ratings , " R: ", "")
expression_10 = "[A-Z]+\\d{2}"
expression_11 = "[A-Z]+\\d{1}"
Post_ratings  <- str_replace_all(Post_ratings, expression_10, "")
Post_ratings  <- str_replace_all(Post_ratings, expression_11, "")
Post_ratings <- str_trim(Post_ratings)
head(Post_ratings,10)

```

```{r results = 'asis'}
#convert all appliable data to numerics 
meta_data <-data.frame (USCF_ID ,Pre_ratings, Post_ratings)
meta_data <- meta_data[-1 ,]
meta_data$USCF_ID <-as.numeric(as.character(meta_data$USCF_ID))
meta_data$Pre_ratings <- as.numeric(as.character(meta_data$Pre_ratings))
meta_data$Post_ratings <- as.numeric(as.character(meta_data$Post_ratings))

knitr::kable(head(meta_data,10),  caption = "Meta Data Table")
```
```{r results = 'asis'}
#combined actionable elements from the odds and evens dataframe  
consolidated_df <- data.frame(reformatted_evens_df,
							  odds_data[,c("Num")], 
							  meta_data)

knitr::kable(head(consolidated_df,10),  caption = "Consolidated Table")
```

```{r results = 'asis'}
#create a reference table to join on the columns
ref_table <-  data.table(consolidated_df[,c("Pair", "Pre_ratings")]) 
colnames(ref_table)[colnames(ref_table)=="Pair"] <- "opp_1" 
ref_table$opp_1 <- as.numeric(as.character(ref_table$opp_1))

knitr::kable(head(ref_table,10),  caption = "Mapping Table")

```

```{r}
#formatting the columns to numerics 
consolidated_df$opp_1 <- as.numeric(as.character(consolidated_df$opp_1))
consolidated_df$opp_2 <- as.numeric(as.character(consolidated_df$opp_2))
consolidated_df$opp_3 <- as.numeric(as.character(consolidated_df$opp_3))
consolidated_df$opp_4 <- as.numeric(as.character(consolidated_df$opp_4))
consolidated_df$opp_5 <- as.numeric(as.character(consolidated_df$opp_5))
consolidated_df$opp_6 <- as.numeric(as.character(consolidated_df$opp_6))
consolidated_df$opp_7 <- as.numeric(as.character(consolidated_df$opp_7))

consolidated_df$opp_1[is.na(consolidated_df$opp_1)] <- 0
consolidated_df$opp_2[is.na(consolidated_df$opp_2)] <- 0
consolidated_df$opp_3[is.na(consolidated_df$opp_3)] <- 0
consolidated_df$opp_4[is.na(consolidated_df$opp_4)] <- 0
consolidated_df$opp_5[is.na(consolidated_df$opp_5)] <- 0
consolidated_df$opp_6[is.na(consolidated_df$opp_6)] <- 0
consolidated_df$opp_7[is.na(consolidated_df$opp_7)] <- 0
```

```{r}
#get all of the pre ratings from the reference table, merge onto the original
merged_df <- merge(x = consolidated_df, y = ref_table, by = "opp_1", all.x = TRUE)
colnames(merged_df)[colnames(merged_df)=="Pre_ratings.y"] <- "opp_1_rating"
colnames(ref_table)[colnames(ref_table)=="opp_1"] <- "opp_2" 
merged_df <- merge(x = merged_df, y = ref_table, by = "opp_2", all.x = TRUE)
colnames(merged_df)[colnames(merged_df)=="Pre_ratings"] <- "opp_2_rating"
colnames(ref_table)[colnames(ref_table)=="opp_2"] <- "opp_3" 
merged_df <- merge(x = merged_df, y = ref_table, by = "opp_3", all.x = TRUE)
colnames(merged_df)[colnames(merged_df)=="Pre_ratings"] <- "opp_3_rating"
colnames(ref_table)[colnames(ref_table)=="opp_3"] <- "opp_4" 
merged_df <- merge(x = merged_df, y = ref_table, by = "opp_4", all.x = TRUE)
colnames(merged_df)[colnames(merged_df)=="Pre_ratings"] <- "opp_4_rating"
colnames(ref_table)[colnames(ref_table)=="opp_4"] <- "opp_5"
merged_df <- merge(x = merged_df, y = ref_table, by = "opp_5", all.x = TRUE)
colnames(merged_df)[colnames(merged_df)=="Pre_ratings"] <- "opp_5_rating"
colnames(ref_table)[colnames(ref_table)=="opp_5"] <- "opp_6"
merged_df <- merge(x = merged_df, y = ref_table, by = "opp_6", all.x = TRUE)
colnames(merged_df)[colnames(merged_df)=="Pre_ratings"] <- "opp_6_rating"
colnames(ref_table)[colnames(ref_table)=="opp_6"] <- "opp_7"
merged_df <- merge(x = merged_df, y = ref_table, by = "opp_7", all.x = TRUE)
colnames(merged_df)[colnames(merged_df)=="Pre_ratings"] <- "opp_7_rating"
colnames(ref_table)[colnames(ref_table)=="opp_6"] <- "opp_7"
```
```{r}
#fill in the nulls
merged_df$opp_1_rating[is.na(merged_df$opp_1_rating)] <- 0
merged_df$opp_2_rating[is.na(merged_df$opp_2_rating)] <- 0
merged_df$opp_3_rating[is.na(merged_df$opp_3_rating)] <- 0
merged_df$opp_4_rating[is.na(merged_df$opp_4_rating)] <- 0
merged_df$opp_5_rating[is.na(merged_df$opp_5_rating)] <- 0
merged_df$opp_6_rating[is.na(merged_df$opp_6_rating)] <- 0
merged_df$opp_7_rating[is.na(merged_df$opp_7_rating)] <- 0

```




```{r}
merged_df$sum_toal_pre_ratings <- (merged_df$opp_1_rating+ merged_df$opp_2_rating+
						merged_df$opp_3_rating+ merged_df$opp_4_rating+
									 merged_df$opp_5_rating+ merged_df$opp_6_rating+
									 merged_df$opp_7_rating)
```


```{r}
merged_df$div<- 7 -(rowSums(merged_df[,23:29] == 0))
merged_df$average_opp_pre_ratings <- merged_df$sum_toal_pre_ratings/merged_df$div

```
###Final Table
Format: Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents 
```{r results = 'asis'}
#final Dataframe
final_df <- merged_df[,c("Pair","Player_Name","odds_data...c..Num...", "Total","Pre_ratings.x", "average_opp_pre_ratings" )]
final_df<- final_df[order(final_df$Pair),] 
#renaming some columns
setnames(final_df, old = c('Pair', 'odds_data...c..Num...','Total','Pre_ratings.x', 'average_opp_pre_ratings' ), 
         new = c('Ranking','Player State','Total Number of Points', 'Player Pre-Ranking','Average Opponent Pre-Ranking'))

knitr::kable(head(final_df ,10),  caption = "Final-Cleaned DataFrame")
```








