---
title: "Week 7 Assignment"
author: "Meaghan Burke"
date: "March 17, 2018"
output: html_document
---


```{r results='hide', message=FALSE, warning=FALSE}
library(knitr)
library(XML)
library(dplyr)
library(tidyr)
library(xml2)
library(rjson)
library(jsonlite)
```

## R Markdown

Pick three of your favorite books on one of your favorite subjects. At least one of the books should have more
than one author. For each book, include the title, authors, and two or three other attributes that you find
interesting.
Take the information that you’ve selected about these three books, and separately create three files which
store the book’s information in HTML (using an html table), XML, and JSON formats (e.g. “books.html”,
“books.xml”, and “books.json”). To help you better understand the different file structures, I’d prefer that you
create each of these files “by hand” unless you’re already very comfortable with the file formats.
Write R code, using your packages of choice, to load the information from each of the three sources into
separate R data frames. Are the three data frames identical?



## XML Version- Jane Austin Novels 
```{r}
# Parse the XML file
xml_novel <- xmlTreeParse("C:/Data 607/Week7 HW/xml_JaneAustin.xml")

#check if R is recognizing the file as a xml file 
class(xml_novel)
```

```{r}
#Access the top node, view the contents & formatting 
top_node_novel <- xmlRoot(xml_novel)
top_node_novel 
```
```{r}
# To put the xml data in a data frame, I extracted the XML values using the function xmlSApply():
top_node_novel  <- xmlSApply(top_node_novel ,
                    function(x) xmlSApply(x, xmlValue))
top_node_novel
```

```{r,  results = 'asis', message=FALSE, warning=FALSE}
#Converted the top node object into a dataframe using data.frame. Formatting required dplyr.
xml_novel_df <- data.frame(t(top_node_novel),
                     row.names=NULL)%>% 
                      separate(Authors, c("Author_1", "Author_2"), "\\|", extra = "merge")%>% 
                      separate(Main_Characters, c("Character_1", "Character_2"), "\\|", extra = "merge")%>%
                      replace(., is.na(.), "")
kable(xml_novel_df,  caption = "XML Jane Austin Dataframe")

```

##HTML Novels-Jane Austin
```{r}
#View the contents & formatting of the HTML object 
html_novel <- "C:/Data 607/Week7 HW/html_JaneAustin_v2.html"
html_novel<- htmlTreeParse(html_novel , error=function(...){}, useInternalNodes = TRUE)
html_novel
```


```{r,  results = 'asis'}
#Converted the HTML object into a dataframe. Converted all the factor columns into characters & renamed the columns. I used the dplyr method. 
html_novel_df <-data.frame(readHTMLTable(html_novel))%>%
  mutate_if(is.factor,as.character)%>%
  rename(Novel = NULL.V1, Author_1 = NULL.V2, Author_2 = NULL.V3, Year_Relased = NULL.V4, Pages =NULL.V5, 
Genre = NULL.V6, Main_Character1 =NULL.V7, Main_Character2 =NULL.V8)

kable(html_novel_df,  caption = "HTML Jane Austin Dataframe")

```



##JSON Novels-Jane Austin

```{r,  results = 'asis'}
#view the contents/formatting of the JSON object 
json_novel_file <- "C:/Data 607/Week7 HW/json_JaneAustin.json.txt"
json_data <- fromJSON(json_novel_file)
json_data 
```
```{r,  results = 'asis'}
#Converted the json object into a dataframe object.  
json_novel_file <- "C:/Data 607/Week7 HW/json_JaneAustin.json.txt"
json_novel_file <- data.frame(fromJSON(json_novel_file  , flatten=TRUE))%>%
  mutate_if(is.factor,as.character)

kable(json_novel_file,  caption = "JSON Jane Austin Dataframe")

```

My Findings: while the three objects store the same information, the ingestion into R required different amounts of effort. For me, JSON was the easiest and most straightforward. 

