---
title: "Week 3 Assignment"
author: "Meaghan"
date: "February 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 3:
```{r, warning=FALSE}
library(stringr)

```

Copy the introductory example. The vector name stores the extracted names:
[1] "Moe Szyslak" "Burns, C. Montgomery" "Rev. Timothy Lovejoy"
[4] "Ned Flanders" "Simpson, Homer" "Dr. Julius Hibbert"


```{r}
raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
names <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
print(names)
```

(A) Use the tools of this chapter to rearrange the vector so that all elements conform to the standard first_name last_name.

```{r}
#remove initials 
names2 <- str_replace(names , "C. ", "")
names2 <- str_replace(names2 , "Rev. ", "")
names2 <- str_replace(names2 , "Dr. ", "")


#Use string split to split the names with comma values. The paste function served as my concatenator. 
names2[2] <- str_split(names2[2], pattern = ",")
names2[2] <- paste(names2[2][[1]][2], names2[2][[1]][1], sep=" ")
names2[5] <- str_split(names2[5], pattern = ",")
names2[5] <- paste(names2[5][[1]][2], names2[5][[1]][1], sep=" ")


print(names2)

```
 (B) Construct a logical vector indicating whether a character has a title (i.e., Rev. and Dr.). 
```{r}

dr_title <- str_detect(names, "Dr. ")
rev_title <- str_detect(names, "Rev. ")
title_df <- data.frame(names,dr_title, rev_title )
title_df
```
(c) Construct a logical vector indicating whether a character has a second name (Middle Name?).
```{r}
middle_name <- str_detect(names, "C. ")
midle_name_df <- data.frame(names,middle_name )
midle_name_df 
```

## Question 4: 
Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.

```{r}
expression_1 = "[0-9]+\\$"

#expression_1 applies to one or more numbers, followed by the "$" sign.

test_1 <- c("5627$", "65666$", "2$")
str_detect(test_1,expression_1)

```
```{r}
expression_2 <- "\\b[a-z]{1,4}\\b"
#expression_2 applies to words with one to four letters
test_2<- c("Meaghan", "test", "regx", "questions")
str_detect(test_2, expression_2)

```
```{r}
expression_3 <- ".*?\\.txt$"
#expression_3 represents a string with a ".txt" at the end. 
test_3 <- c("file.txt","file.pdf","file.exe","homework.txt")

str_detect(test_3, expression_3)
```
```{r}
expression_4<-"\\d{2}/\\d{2}/\\d{4}"
#expression of numbers in the xx/xx/xxxx, relavent for dates.

test_4 <-c("05/26/1993", "05/10/2015", "12/31/2017")
str_detect(test_4, expression_4)
```
```{r}
expression_5 <- "<(.+?)>.+?</\\1>"
#this expression is a tag format.The first part is an expression1 inside the "<>" followed by expression2 outside "<>", followed by expression1 inside "</>"
test_5 <- c("<test1>test2</test1>", "<html>blue</html>")
str_detect(test_5, expression_5)
			
```

#Question 9

The following code hides a secret message. Crack it with R and regular expressions. Hint: Some of the characters are more revealing than others! The code snippet is also available in the materials at www.r-datacollection.com.



```{r}
extra_credit <- paste("clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5 fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr")

upper_case <- unlist(str_extract_all(extra_credit, "[[:upper:].]{1,}"))
upper_case

length(upper_case)
upper_case <- paste(upper_case, sep="", collapse="")
secret_message <- str_replace_all(upper_case, "[.]", " ")

secret_message
```

